---
title: "Data Science for Business Project Part 1"
subtitle: Section C Team 59Esrat Esha, Santiago de la Flor Giuffra, Shrirang Ojha,
  Yining Zhang, Zoe Li
output: html_notebook
---

#1. Importing libraries

```{r}
#We added some functions to the files below
source("DataAnalyticsFunctions.R")
source("PerformanceCurves.R")

library(tree)
library(tidyverse)
library(tibble)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2) 
library(corrplot)
library(RColorBrewer)
library(quantreg)
library(car)
library(mice)
library(glmnet)
library(tree)
library(partykit)
library(randomForest)
library(missForest)
library(VIM)
```

#2. Data Understanding

##2.1.Loading Data Dictonary

```{r}
#We added some functions to the files below

data_dictionary<-read.csv("Data_Dictionary.csv")
data_dictionary
```

##2.2.Loading the main dataset

```{r}
#As the compettion in Kaggle already closed, we only have access to the train dataset.
#We will split it into train and test on further steps.
dataset<- read_csv('Train_Dataset.csv')
dataset
```

##2.3.Column exploration

```{r}
str(dataset)
```

#3. Data Preparation

##3.1 Transformation of char columns

```{r}
#From str(dataset),we can notice below that there is a few columns that are being considered as characters that need to be transformed to integers or floats. 

dataset <- dataset %>%
  mutate(
    Population_Region_Relative = as.numeric(Population_Region_Relative),
    Age_Days = as.integer(Age_Days),
    Employed_Days = as.integer(Employed_Days),
    Registration_Days = as.integer(Registration_Days),
    ID_Days = as.integer(ID_Days)
  )

#As the warnings suggest, we have introduced NA's to the columns. On a later step, we will focus on addressing the NA's for this and other columns.
```

##3.2 First check of NA values

```{r}
na_percentage <- dataset %>%
  summarise(across(everything(), ~ sum(is.na(.)) / length(.) * 100)) %>%
  pivot_longer(cols = everything(), names_to = "Column", values_to = "NA_Percentage") %>%
  arrange(desc(NA_Percentage))

print(na_percentage)
```

##3.2 Eliminating unnecesary columns

```{r}
#The columns below will be dropped because they do not hold importance for our analysis and also hold NAs.
#We are also dropping columns that have a substancial amount of NAs

#Not relevant for our analysis
dataset <- subset(dataset, select = -Accompany_Client)
dataset <- subset(dataset, select = -Application_Process_Hour)
dataset <- subset(dataset, select = -Application_Process_Day)
dataset <- subset(dataset, select = -Type_Organization)
dataset<-subset(dataset, select= -Score_Source_2)
dataset <- subset(dataset, select = -Mobile_Tag)	

#Substancial amount of NAs
dataset <- subset(dataset, select = -Own_House_Age)
dataset <- subset(dataset, select = -Score_Source_1)
dataset <- subset(dataset, select = -Social_Circle_Default)
dataset <- subset(dataset, select = -Score_Source_3)		
dataset <- subset(dataset, select = -Credit_Bureau)	
```

##3.3 Converting columns without NAs into factors

```{r}
dataset$Client_Permanent_Match_Tag<-as.factor(dataset$Client_Permanent_Match_Tag)
dataset$Client_Contact_Work_Tag<-as.factor(dataset$Client_Contact_Work_Tag)
dataset$Default <- as.factor(dataset$Default)
```

##3.4 Handling NAs for Categorical columns

### a) Strategy 1 for NA: Converting to a Did not disclose factor

####Client_Income_Type

```{r}
# Initial column repetition
print("Initial columns")
table(dataset$Client_Income_Type, useNA = "ifany")

# Replac NA values with "Income not Disclosed"
dataset$Client_Income_Type <- ifelse(is.na(dataset$Client_Income_Type), "Income not Disclosed", dataset$Client_Income_Type)

# These are the four categories we want to keep
categories_to_keep <- c("Commercial", "Service", "Retired", "Govt Job")

# Replace other categories with "Other or Income not Disclosed"
dataset$Client_Income_Type <- ifelse(dataset$Client_Income_Type %in% categories_to_keep, 
                                     dataset$Client_Income_Type, 
                                     "Other or Income not Disclosed")

# # Convert the column to factor
dataset$Client_Income_Type <- as.factor(dataset$Client_Income_Type)

# New column repetition
print("New columns")
table(dataset$Client_Income_Type, useNA = "ifany")
```

####Client_Education

```{r}
## Initial column repetition
print("Initial columns")
table(dataset$Client_Education, useNA = "ifany")

# Replace NA values with "Client Education Not Disclosed"
dataset$Client_Education <- ifelse(is.na(dataset$Client_Education), "Client Education Not Disclosed", dataset$Client_Education)
# Combine "Graduation" and "Post Grad" into one category: "Grad and Post Grad"
dataset$Client_Education <- ifelse(dataset$Client_Education %in% c("Graduation", "Post Grad"), 
                                   "Grad and Post Grad", 
                                   dataset$Client_Education)

# Convert the column to factor
dataset$Client_Education <- as.factor(dataset$Client_Education)

# New column repetition
print("New columns")
table(dataset$Client_Education, useNA = "ifany")
```

####Marital Status

```{r}
## Initial column repetition
print("Intial columns")
table(dataset$Client_Marital_Status, useNA = "ifany")

# Replace NA values with "Marital Status Not Disclosed"
dataset$Client_Marital_Status <- ifelse(is.na(dataset$Client_Marital_Status), "Marital Status Not Disclosed", dataset$Client_Marital_Status)

# Convert the column to factor
dataset$Client_Marital_Status <- as.factor(dataset$Client_Marital_Status)

# New column repetition
print("New columns")
table(dataset$Client_Marital_Status, useNA = "ifany")
```

####Client_Housing_Type

```{r}
## Initial column repetition
print("Intial columns")
table(dataset$Client_Housing_Type, useNA = "ifany")

# Replace NA values with "Housing Type Not Disclosed"
dataset$Client_Housing_Type <- ifelse(is.na(dataset$Client_Housing_Type), "Housing Type Not Disclosed", dataset$Client_Housing_Type)

# Convert the column to factor
dataset$Client_Housing_Type <- as.factor(dataset$Client_Housing_Type)

# New column repetition
print("New columns")
table(dataset$Client_Housing_Type, useNA = "ifany")
```

###Client_Occupation We will replace NA values with Occupation_Type_Not_Disclosed dummy

```{r}
## Initial column repetition
print("Intial columns")
table(dataset$Client_Occupation, useNA = "ifany")

# Replace NA values with "Occupation Type Not Disclosed"
dataset$Client_Occupation <- ifelse(is.na(dataset$Client_Occupation), "Occupation Type Not Disclosed", dataset$Client_Occupation)

#Keeping on
frequency_table<-data.frame(table(dataset$Client_Occupation)) %>%  
  filter(Freq>2000)
categories_to_keep <- c(frequency_table$Var1) 

# Replace other categories with "Other"
dataset$Client_Occupation <- ifelse(dataset$Client_Occupation %in% categories_to_keep, 
                                     dataset$Client_Occupation, 
                                     "Other")
# New column repetition
print("New columns")
table(dataset$Client_Occupation, useNA = "ifany")
```

### b) Strategy 2 for NA: Dropping the records

####Client_Gender

```{r}
#We have signifcantly less NA values here, therefore, we decide to drop the records

# Initial column repetition
print("Intial columns")
table(dataset$Client_Gender, useNA = "ifany")

#Removing XNA and Client Gender
dataset <- subset(dataset, Client_Gender != 'XNA')
dataset <- subset(dataset, !is.na(Client_Gender))

#Converting the columns into factors
dataset$Client_Gender <- as.factor(dataset$Client_Gender)

# New column repetition
print("New columns")
table(dataset$Client_Gender, useNA = "ifany")
```

###c) Strategy 3 for NA: Replacing with mode

####Loan_Contract_Type

```{r}
#As the loan contract type has to be disclosed, we will replace the NA values with the mode

# Initial column repetition
print("Intial columns")
table(dataset$Loan_Contract_Type, useNA = "ifany")

#Converting NAs into CL
dataset$Loan_Contract_Type <- ifelse(is.na(dataset$Loan_Contract_Type), "CL", dataset$Loan_Contract_Type)

#Converting the columns into factor
dataset$Loan_Contract_Type <- as.factor(dataset$Loan_Contract_Type)

# New column repetition
print("New columns")
table(dataset$Loan_Contract_Type, useNA = "ifany")
```

##3.5. Data Preparation for Numerical columns

```{r}
#For the numerical values (Less than 5% of NAs), we took a different approach and used mice with a car with 10 iterations
No_NA_dfs <- mice(dataset,m=1,maxit=10,meth='cart',seed=300)
```

##3.6. Saving the new clean dataframe

```{r}
complete(No_NA_dfs, 1)
write.csv(complete(No_NA_dfs, 1), "Clean_data.csv")
```
